Choose a License

### 2024-09-21 choose_license.qmd
A few improvements to license infos

Whenever you create a literary or artistic work (such as a text, image, video, or software), the copyright law in most countries limits other people from copying, modifying, and sharing it without your express permission. [If the work was created as part of your job, it might be your employer who holds the copyright to the work, depending on the country and contract.]{.aside} This even applies if you make it available to others (e.g., on your website): First and foremost they are not allowed to copy, modify, or share it. This legal default of "all rights reserved" runs counter to many cultural and scientific processes. Copyright licenses enable authors to free up their works for reuse by others.

@@

When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license that goes further than the open access license available to readers of the journal [@Rumsey2022]. However, technically this is not required for publishers to provide their services [@Suber2022] and authors should oppose this practice. If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.

@@

Many boilerplate licenses are available to apply to your work. Which license is appropriate depends on several factors, including existing licenses in place and the type of work, but also your personal considerations. We strongly recommend to apply a _free/open_ license to your work, which means that the work "can be freely studied, applied, copied and/or modified, by anyone, for any purpose" [@Moller2015]. Importantly, this also means that others do not need to ask or notify the author and that they can use it for commercial purposes. By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that derivative works are only shared under the same license, and that no further restrictions are imposed on the work. Software licenses may additionally require to make the source code available to everybody the software is shared with. Because there are many free/open licenses available, the licenses discussed here only represent a recommended subset.

@@

::: {#tip-dual-licensing .callout-tip}
#### Dual-licensing

@@

One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to keep the license notice intact]

One solution is to make such a work simultaneously available under two licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video]. This is called dual-licensing and makes it easier to reuse both the rendered document as well as the code. For example, one could write:

@@

As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use. There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015], but we would like to highlight one reason in particular.

@@

The answer is that you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project). The easiest approach is to make note of every foreign work included in your project and record its license. Then, dual-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.

@@

For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022].
:::

>>>>

Whenever you create a literary or artistic work (such as a text, image, video, or software), the copyright law in most countries limits other people from copying, modifying, and sharing it without your express permission. [If the work was created as part of your job, it might be your employer who holds the copyright to the work, depending on the country and contract.]{.aside} This even applies if you make it available to others (e.g., on your website): First and foremost they are not allowed to copy, modify, or share it. This legal default of "all rights reserved" was created to benefit publishers, not authors, [@Fogel2006] and runs counter to many cultural and scientific processes. Copyright licenses enable authors to free up their works for reuse by others.

@@

When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment. This would give them more rights than the readers of the article have through the chosen open access license [@Rumsey2022], and exceeds exceeds by far what is necessary to make publication possible [@Suber2022]. Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has. If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.

@@

Many boilerplate licenses are available to apply to your work. Which license is appropriate depends on several factors, including existing licenses in place and the type of work, but also your personal considerations. We strongly recommend to apply a _free/open_ license to your work, which means that the work "can be freely studied, applied, copied and/or modified, by anyone, for any purpose" [@Moller2015]. Importantly, this also means that others do not need to ask or notify the author and that they can use it for commercial purposes. By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that derivative works are only shared under the same license, and that no further restrictions are imposed on the work. Software licenses may additionally require to make the source code available to everybody the software is shared with and often require to display the full text of the license upon usage. Because there are many free/open licenses available, the licenses discussed here only represent a recommended subset.

@@

::: {#tip-multi-licensing .callout-tip}
#### Multi-licensing

@@

One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to display the full text of the license]

One solution is to make such a work simultaneously available under two (or more) licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video]. This is called multi-licensing and makes it easier to reuse both the rendered document as well as the code. For example, one could write:

@@

As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use. There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023], but we would like to highlight one reason in particular.

@@

The answer is that you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project). The easiest approach is to make note of every foreign work included in your project and record its license. Then, multi-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.

@@

For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022Version].
:::
